---
title: "redland: create, query and write RDF graphs"
author: "Peter Slaughter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{redland RDF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8} 
---

## Introduction

The `redland` R package provides methods to create, query and write to disk data stored in the Resource Description Framework (RDF). RDF provides a standardized way to make statements about resources and the relationships between them. Typical resources include datasets that are available via URLs. Resources don't have to be available online, for example a resource could also be the name of the person that created a dataset. A collection of RDF statements comprise an RDF graph, which can be read and interpreted by an RDF capable software application that uses the resources described in the graph. 

An introduction to RDF can be found at http://www.w3.org/TR/2014/NOTE-rdf11-primer-20140624.

## Examples

The following example reads an RDF graph from a disk file that was previously saved in the RDF/XML format:

```{r}
library(redland)
world <- new("World")
storage <- new("Storage", world, "hashes", name="", options="hash-type='memory'")
model <- new("Model", world=world, storage, options="")
parser <- new("Parser", world)
parseFileIntoModel(parser, world, system.file("extdata", "dc.rdf", package="redland"), model)
  
```
Next the RDF graph is queried for statements of interest, using the [SPARQL query syntax](http://www.w3.org/TR/rdf-sparql-query)

```{r}
queryString <- 'PREFIX dc: <http://purl.org/dc/elements/1.1/> SELECT ?a ?c WHERE { ?a dc:creator ?c . }'
query <- new("Query", world, queryString, base_uri=NULL, query_language="sparql", query_uri=NULL)
queryResult <- executeQuery(query, model)
results <- getResults(query, model, "rdfxml")

cat(sprintf("Results from query: %s\n", results))
```

Next, additional statements can be added to the RDF graph:
```{r}
stmt <- new("Statement", world=world, 
        subject="http://www.dajobe.org/",
        predicate="http://purl.org/dc/elements/1.1/language",
        object="en")
addStatement(model, stmt)
```

Now the RDF graph can be written to disk:

```{r}
serializer <- new("Serializer", world, mimeType="application/rdf+xml")
status <- setNameSpace(serializer, world, namespace="http://purl.org/dc/elements/1.1/", prefix="dc")  
filePath <- tempfile(pattern = "file", tmpdir = tempdir(), fileext = ".rdf")
status <- serializeToFile(serializer, world, model, filePath)
```

